---
output:
  pdf_document: default
  html_document: default
---
**Stat 230: Linear Models**  
**Homework 3**  
**Professor Ding**  
**Lev Golod**  


```{r prelim, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
```


#### Question 3, part A
```{r q3a}
## Input the data 
x1<-c(2.34,2.46,2.83,2.04,2.69)
x2<-c(2.64,3.00,3.19,3.83)
x3<-c(2.61,2.07,2.80,2.58,2.98,2.30)
x4<-c(1.32,1.62,1.92,0.88,1.50,1.30)
x5<-c(0.41,0.83,0.53,0.32,1.62)
mylist <- list(x1,x2,x3,x4,x5)
names(mylist) <- c('coat1', 'coat2', 'coat3', 'shirt1', 'shirt2')
# mylist


## Identify mean[j] and n[j] for each treatment grp
means <- sapply(mylist, mean)
ns <- sapply(mylist, length)
             
## Identify grand mean and total n, total J             
ybar <- mean(unlist(mylist))
n <- length(unlist(mylist))
J <- length(mylist)


## Sum of Squares - Regression 
# sum(sapply(mylist, function(x) 
#   length(x) * (mean(x) - ybar)^2)) == sum(ns * (means - ybar)**2)
df_reg <- (J-1)
SSR <- sum(ns * (means - ybar)**2)
MSR <- SSR/df_reg


## Sum of Squaress - Error 
SSE <- sum(sapply(mylist, function(x) 
  sum( (x - mean(x))**2 )))
df_err <- (n-J)
MSE <- SSE/df_err


## Sum of Squares - Total (just to check SST = SSE + SSR)
SST <- sum( (unlist(mylist)-ybar)**2 )
df_tot <- (n-1)
# df_tot == df_err + df_reg
# abs(SST - SSE - SSR) < 1e-10


## Calculate F statistic & p value
Fstat <- MSR/MSE
p <- pf(Fstat, J-1, n-J, lower.tail = FALSE)


## Display a table with ANOVA results
names <- c('DF', 'SS', 'MS', 'F', 'P')
source <- c("Regression", 'Error', 'Total')
df <- c(df_reg, df_err, df_tot)
SS <- c(SSR, SSE, SST)
MS <- c(MSR, MSE, NA)
F  <- c(Fstat, NA, NA)
# P  <- c(p, NA, NA)
P  <- c(p, NA, NA)
q3a_anova <- cbind(df, SS, MS, F, P)
rownames(q3a_anova) <- source
colnames(q3a_anova) <- names
q3a_anova_display <- apply(q3a_anova, c(1,2), function(x)
  as.character(round(x,4)))
q3a_anova_display['Regression','P'] <- format(p)
kab <- kable(q3a_anova_display, digits = 5)
cat("Question 3, Part A: Anova Table")
```

```{r q3a_kab, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kab
```

The p-value is small, so we reject the Null Hypothesis.  
H0: Beta_1 = ... Beta_J  
We conclude that at least some of the pairs are different from one another.  



```{r q3a_check}
## Compare with R's built in fb 
# put data into a data frame 
labs <- mapply(rep, names(mylist), sapply(mylist, length))
dat_q3a <- data.frame(
  time = unlist(mylist, use.names = FALSE),
  brand = unlist(labs, use.names = FALSE)
)
q3a_lm <- lm(time ~ brand, data = dat_q3a)
# # Check that my result matches R's
# abs(q3a_anova[1:2,] - anova(q3a_lm)) < 1e-10
anova(q3a_lm)
```



#### Question 3, part B - this approach uses contrasts 
```{r q3b}
# Lambda
L <- c(ns[1:3]/sum(ns[1:3]), -ns[4:5]/sum(ns[4:5]))
# abs(sum(L)) < 1e-10

# A function to output a t value and a TWO-SIDED p value for a contrast
contr <- function(L, MSE, means, ns, df){
  tstat <- abs(sum( means*L )) / (
    ( MSE * sum((L**2) / ns) )**0.5
  )
  p <- pt(tstat, df, lower.tail = FALSE)
  c(tstat = tstat, pval = p)
}
# cut the p-val in half since this is a 1 sided test
contr(L, MSE, means, ns, n-J)["tstat"]; contr(L, MSE, means, ns, n-J)["pval"]/2
```

The p value is small so we reject the Null Hypothesis.  We conclude these brands  
of coats are sturdier than these brands of shirts.  



### Question 3, Part C 
```{r q3c}
checkeq <- function(a, b = 1e-10) abs(a) < b

# Find contrasts 
# I found these contrasts using Mathematica
v1 <- c(119/11, -(130/11), 0, 0, 1)
v2 <- c(-(4643/1012), -(8329/2024), 15591/2024, 0, 1)
v3 <- c(-(427/14198), 1621/28396, -(3329/28396), -(12917/14198), 1)
v <- list(v1,v2,v3)

# # All must sum to 0
# sapply(v, function(x) checkeq(sum(x)))
# 
# # All must be orthogonal to L
# sapply(v, function(x) checkeq(t(x) %*% L ) )
# 
# # ALl must be mutually orthogonal - dot product = 0
# t(v1) %*% L
# t(v2) %*% L
# t(v3) %*% L
# t(v1) %*% v2
# t(v1) %*% v3
# t(v2) %*% v3
cons <- c(list(L), v)
names(cons) <- c('original', 'orthog1', 'orthog2', 'orthog3')
# cons

# Find the Sum of Squares for each contrast 
# This is defined on Page 124 of 505 of the Christiansen book
sscon <- function(L, means, ns) sum(L*means)**2 / sum( (L**2)/ns )
sapply(cons, sscon, means = means, ns = ns)
```



#### Question 3, Part D
```{r q3d}
# Find a CI for the difference between shirt1 and shirt2
# Identify the contrast that corresponds to this test
L2 <- c(0,0,0,1,-1)

# t critical value 
alpha <- 0.05
tstar <- qt(1-alpha/2, n-J)

# center of the CI
center <- t(L2) %*% means

# SD of the test statistic 
sd <- sqrt( MSE * sum((L2**2)/ns) )

# Display the CI 
center + c(-1,1) * sd * tstar
```

Since the CI does not contain zero, we conclude that shirt1 is sturdier.  
