**Stat 230: Linear Models**  
**Homework 2**  
**Professor Ding**  
**Lev Golod**  

-----------

```{r prelim, include=FALSE}
library(ggplot2)
library(rgl)
load('./family.rda')
names(family) <- tolower(names(family))
```

#### Question 5: Part A
```{r Q5_PartA}
# X matrix has column of 1s (for intercent), weight, and BMI
X <- as.matrix(family)[,c(5,6)]
X <- apply(X, 2, as.numeric)
X <- cbind(rep(1, nrow(X)), X)

# Y vector has height which is what we are predicting
Y <- as.matrix(family)[,4,drop = FALSE]
Y <- apply(Y, 2, as.numeric)

# Beta-hat = (X'X)^-1 X' Y
betahat = solve(t(X) %*% X) %*% t(X) %*% Y

# Residuals : e-hat = Y - Yhat = Y - X beta-hat
Yhat <- X %*% betahat
ehat = Y - Yhat
colnames(ehat) <- NULL


print(round(betahat,3))
print(t(round(ehat,3)))
```


#### Question 5: Part B
```{r Q5_partB}
myplot <- ggplot(family, aes(x = weight, y = height)) +
  geom_point(aes(size = bmi)) +
  theme(legend.position = 'none') +
  ggtitle("Question 5B: Plot regression of height ~ weight", 
          subtitle = "Point size corresponds to bmi")
myplot + geom_smooth(method = 'lm', se = FALSE)
```
  
Comment: We see that most of the larger points are below the regression line.  
For a given weight, higher bmi individuals tend to be shorter than predicted,  
and lower bmi individuals tend to be taller than predicted. This explains the  
fact that the coefficient for bmi is negative: it means that, holding all else  
constant, the relationship bewtween height and bmi is engative.  


#### Question 5: Part C
```{r Q5_partC}
# open3d()
plot3d(x = X[,'weight'], y = X[,'bmi'], z = Y,
       xlab = "weight", ylab = "bmi", zlab = "height")
a <- betahat['weight',]
b <- betahat['bmi',]
c <- -1
d <- betahat[1,]
planes3d(a,b,c,d)
rgl.snapshot("HW2_Q5_partC_3dplot.png")
```



